---
title: "GWHAP PART 2 - HAPLOTYPES BLOCK"
author: "SK"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GWHAP PART 2 - HAPLOTYPES BLOCK}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  error=FALSE,
  message=FALSE,
  warning=FALSE,
  comment = "#>"
)
```
```{r}
library(gwhap)
library(data.table)
library(readr)
```
# Genome-Wide HAPloptype analysis  
## PART 2 : Haplotype blocks
### GWHAP will determine haplotypes blocks where neigboring markers are less that \delta  apopart on the map (in cM)  

See vignette "Part1-Maps" for computing genetic maps of markers

### Haplotype blocks  are defined on  genetic maps :

TODO unify output format from create_ and get_
```{r}
snp_physical_positions = system.file("extdata", "example_high_density.bim", package="gwhap", mustWork=TRUE)
snp_list=get_bim_file(snp_physical_positions)
genetic_map_dir=dirname(snp_physical_positions)

filepaths = gwhap:::gwhapConfig$genmap_toy_interpolated_1000$filepaths
encodings = gwhap:::gwhapConfig$genmap_toy_interpolated_1000$encodings

# get an instance of Genetic_Map and readData
interp1000_genetic_map = Genetic_Map(filepaths=filepaths, encodings=encodings)
interp1000_genetic_map = readData(interp1000_genetic_map)


genetic_map=create_augmented_genetic_map(snp_physical_positions=snp_physical_positions, 
                                        genetic_map=interp1000_genetic_map,
                                         genetic_map_dir=genetic_map_dir, 
                                         map_name='1000_genome_interpolated', 
                                         save_genetic_map=FALSE)
tmpDir = tempdir()
create_augmented_genetic_map(snp_physical_positions=snp_physical_positions, 
                                          genetic_map=interp1000_genetic_map,
                                         genetic_map_dir=genetic_map_dir, 
                                         map_name='1000_genome_interpolated', 
                                         save_genetic_map=TRUE,
                                         output = tmpDir)
augmented_map_df = get_augmented_genetic_map(augmented_genetic_map_dir= tmpDir,
                                             chromosomes=1:23)
head(genetic_map)
head(augmented_map_df)
```




* Create blocs with a single delta

TODO : error when no blocks were created on a given chromosome

```{r}
#create_blocs(genetics_map=augmented_map_df, 
#             delta=1e-2,
#             save_blocs=TRUE,
#             output=tempdir())
#df_blocks=get_blocs(blocs_dir = tempdir(),
#          chromosomes = 1:23)
#head(df_blocks)
```

* Create blocs with multiple values of delta
```{r}
#deltas = c(1e-3, 1e-2)
#dfs_blocks = c()
#for(delta in deltas){
#  df_blocks_tmp = create_blocs(genetics_map=augmented_map_df,
#                               delta=delta,
#                               save_blocs=FALSE)
#  dfs_blocks <- rbind(dfs_blocks, df_blocks_tmp)
#}
#df_blocs=do.call(rbind,dfs_blocks)
#head(df_blocs)
```

* GWHAP provides high-level function to produce images for block distribution

see vignette "visualization"

TODO : rebuild vignette visualization with toy examples
