---
title: "GWHAP PART1 - GENETIC MAPS INTERPOLATION"
author: "SK"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GWHAP PART1 - GENETIC MAPS INTERPOLATION}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  error=FALSE,
  message=FALSE,
  warning=FALSE,
  comment = "#>"
)
```
```{r}
library(gwhap)
library(data.table)
library(readr)
```
# Genome-Wide HAPloptype analysis  
## PART 1 : Genetic Map
### GWHAP needs to depermine the location the genetic map (i.e in cM) of the phased markers. 

For every marker postion in bp, GWHAP will try find it in a reference genetic map.
If the marker's position is not in the genetic map, its position in cM is computed using linear interpolation. 
GWHAP can use several reference maps format :  Rutgers genetic map v3 , 1000 Genomes ...
See vignette "Ressources" for details
TO DO : code that check the download URL for all downloaders

### Reading the the reference genetic map :


* GWHAP can read reference maps in Rutgers Maps filename and column format :
```{r}
genetic_map_file=system.file("extdata", "RUMapv3_B137_chr1.txt",package="gwhap", mustWork=TRUE)
genetic_map=get_rutgers_map(genetic_map_file) # for chr. 1
head(genetic_map)
```


NOT WORKING : GWHAP can read 1000 genomes maps:

TODO : use original 1000 genomes map files

* GWHAP can read  1000 genomes genetic maps file format
```R
genetic_map_file=system.file("extdata", "chr1_1000_Genome.txt",package="gwhap", mustWork=TRUE)
genetic_map=get_1000_genome(genetic_map_file)
head(genetic_map)
```



* GWHAP can read compressed (.gz) 1000 genomes maps using 3 columns [ rsid ; position ; cM]  :
```{r}
genetic_map_file=system.file("extdata", "chr1.interpolated_genetic_map.gz",package="gwhap", mustWork=TRUE)
genetic_map=get_1000_genome_interpolated_map(genetic_map_file)
head(genetic_map)
```




### Reading the list of SNPs to place on the map :

* GWHAP can read a list of SNPs using the bgi/bgen file :
```{r}
snp_file=system.file("extdata","haplotypes.bgen.bgi", package="gwhap", mustWork=TRUE)
snp_list=get_bgi_file(snp_file)
head(snp_list)
```

* GWHAP can read a list of SNPs using any text file in .bim format  :
```{r}
snp_physical_positions = system.file("extdata", "example.bim", package="gwhap", mustWork=TRUE)
snp_list=get_bim_file(snp_physical_positions)

head(snp_list)
```


### GWHAP provide build-in function to interpolate a list of SNPs using a reference map :
TODO fix buggy "interp_in" list : keep snps reference needed


*  Interpolate maps for bim files format  and 1000 Genomes reference maps:
```{r}
genetic_map_file=system.file("extdata", "chr1.interpolated_genetic_map.gz",package="gwhap", mustWork=TRUE)
genetic_map_dir=dirname(genetic_map_file)
snp_physical_positions =snp_physical_positions = system.file("extdata", "example.bim", package="gwhap", mustWork=TRUE)
interpolated_map=create_augmented_genetic_map(snp_physical_positions=snp_physical_positions, genetic_map_dir=genetic_map_dir, map_name='1000_genome_interpolated', save_genetic_map=FALSE)


head(interpolated_map)
```


TODO : Use an ybgen.bgi filenames format (use list?) and  1000 genomes maps

* Using UKB Biobank filenames format (bgen.bgi) and  1000 genomes maps : 

```{r}
genetic_map_file=system.file("extdata", "chr1.interpolated_genetic_map.gz",package="gwhap", mustWork=TRUE)
genetic_map_dir=dirname(genetic_map_file)
snp_file=system.file("extdata","haplotypes.bgen.bgi", package="gwhap", mustWork=TRUE)
snp_physical_positions=sprintf("%s/ukb_",dirname(snp_file))
interpolated_map=create_augmented_genetic_map(snp_physical_positions=snp_physical_positions, genetic_map_dir=genetic_map_dir, map_name='1000_genome_interpolated', save_genetic_map=FALSE)

head(interpolated_map)
```



* Using UKB Biobank filenames format (bgen.bgi) and  Rutgers maps:

```{r}
genetic_map_file=system.file("extdata", "RUMapv3_B137_chr1.txt",package="gwhap", mustWork=TRUE)
genetic_map_dir=dirname(genetic_map_file)
snp_physical_positions="inst/extdata/ukb_"
interpolated_map=create_augmented_genetic_map(snp_physical_positions=snp_physical_positions, genetic_map_dir=genetic_map_dir, map_name='rutgers', save_genetic_map=FALSE)

head(interpolated_map)
```


