---
title: "GWHAP PART 4 : HAPLOTYPE ASSOCIATION TEST"
author: "SK"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GWHAP PART 4 : HAPLOTYPE ASSOCIATION TEST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  error=FALSE,
  message=FALSE,
  warning=FALSE,
  comment = "#>"
)
```
```{r}
library(gwhap)
library(data.table)
library(readr)
```
# Genome-Wide HAPloptype analysis  
## PART 4 : Haplotype Association Test
### GWHAP will compute association test for a phenotype and a given haplotype count matrix

* Get Haplotype count matrix :

```{r}
toy_hap_fn <- system.file("extdata", "haplotypes_many_samples.haps" ,package="gwhap", mustWork=TRUE)
phased_dl.haps = phased_data_loader.haps(toy_hap_fn)
samples_selected = c("sample_0","sample_1","sample_2","sample_3","sample_4","sample_5","sample_6","sample_7")
haplotypes.haps = determine_haplotypes_per_bloc(phased_dl.haps, chromosome="1", 
                                                start=1, 
                                                end=2,
                                                sample_iid=samples_selected,
                                                sample_bgen_iid_code=samples_selected)

# colnames(haplotypes.bgen)=paste("hap", colnames(haplotypes.bgen), sep="_")
haplotypes.haps
```
TODO : fix lm_test for Y matrix (several phenotypes)
WARNING : For N subjects, needs strictly less  than N-1 haplotypes
* Run test on random phenotype:
```{r}

X=data.frame(haplotypes.haps)
Y=data.frame(pheno=rnorm(nrow(X)))
lm_test_haplotypes(X=X, Y=Y,kind='all')
```
* loop to evaluate pvalue distribution and FPR :
```{r}
pv=c()
n=5000
for (i in 1:n) {
X=data.frame(haplotypes.haps)
Y=data.frame(pheno=rnorm(nrow(X)))
lm=lm_test_haplotypes(X=X, Y=Y,kind='bloc')
pv=c(pv,lm$bloc$p_value)
}
y=pv
plot(-log10(1:n/(n+1)),-log10(sort(y)),col="red",pch='o',xlab = "-log10 expected",ylab = "-log10 observed")
abline(a=0,b=1)
mean(pv<0.25)
mean(pv<0.1)
mean(pv<0.05)

```